---
title: "Exploratory analysis"
author: "Ana Abad"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(gtsummary)
library(DataExplorer)
library(ggplot2)
library(ggpubr)

```



```{r}
load("ATP_final.RData")
```

## 1. Summary
```{r}
summary(datos_combinados)
```

## 2. Variable characteristics

### games


```{r}
table(datos_combinados$games, useNA = "always")
```




### tourney_name

```{r}
n_distinct(datos_combinados$tourney_name)
n_distinct(datos_combinados$tourney_id)
```

### Surface
```{r}
unique(datos_combinados$surface)
table(datos_combinados$surface)
```
### tourney_level and level

```{r}

unique(datos_combinados$tourney_level)
table(datos_combinados$tourney_level)

```

### round

```{r}
table(datos_combinados$round)
```

### winner_id, loser_id

```{r}

n_distinct(datos_combinados$winner_id)
n_distinct(datos_combinados$loser_id)
```
### winner/loser_entry

```{r}
unique(datos_combinados$winner_entry)
unique(datos_combinados$loser_entry)

```

### hands 

```{r}

unique(datos_combinados$winner_hand)
unique(datos_combinados$loser_hand)

```

### country

```{r}

n_distinct(datos_combinados$winner_ioc)
n_distinct(datos_combinados$loser_ioc)

```
### Age
```{r}
summary(datos_combinados$winner_age)
summary(datos_combinados$loser_age)
is.integer(datos_combinados$winner_rank)
summary(datos_combinados$dif_age)
```


### Ranking


```{r}

missing_data_summary <- datos_combinados %>%
  summarize(
    winner_rank_missing = sum(is.na(winner_rank)),
    loser_rank_missing = sum(is.na(loser_rank)),
    .by = year
  )


missing_data_long <- missing_data_summary %>%
  pivot_longer(cols = c(winner_rank_missing, loser_rank_missing), 
               names_to = "rank_type", 
               values_to = "missing_count")

missrankbyyear <- ggplot(missing_data_long, aes(x = year, y = missing_count, color = rank_type, group = rank_type)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("winner_rank_missing" = "#000078", "loser_rank_missing" = "#73edff"),
    labels = c("Winner Rank Missing", "Loser Rank Missing")
  ) +
  labs(
    title = "Missing Values in Player Ranks Over Time",
    x = "Year",
    y = "Number of Missing Values",
    color = "Rank Type"
  ) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(
    text = element_text(size = 12),
    legend.position = "top"
  )




ggsave(filename = "Rankbyyear.tif", plot = missrankbyyear, device = "tiff",  dpi = 300)

```


### Summary 


```{r}

rounds <- c(
  "F" = "Final",
  "QF" = "Quarterfinal",
  "R32" = "Preliminary", 
  "R16" = "Preliminary", 
  "R64" = "Preliminary", 
  "R128" = "Preliminary", 
  "BR" = "Preliminary", 
  "RR" = "Preliminary", 
  "ER" = "Preliminary"
)

datos_combinados <- datos_combinados %>%
  mutate(round_category = recode(round, !!!rounds))

tbl_summary(data = datos_combinados, include = c(surface, tourney_level, winner_hand, loser_hand, dif_age, winner_rank, loser_rank, dif_rank, round_category, best_of, games))
```




## 3. Report by PR and LL

```{r}


tbl_summary(data = datos_combinados, by = PR, include = c(year, surface, tourney_level, winner_hand, loser_hand, dif_age, winner_rank, loser_rank, dif_rank, best_of, round_category, games))


tbl_summary(data = datos_combinados, by = LL, include = c(year, surface, tourney_level, winner_hand, loser_hand, dif_age, winner_rank, loser_rank, dif_rank, best_of, round_category, games))




```


## 4. Distributions



```{r}

plot_histogram(datos_combinados)
plot_bar(datos_combinados)


```

### Normality in continuous variables

```{r}
ggqqplot(datos_combinados$dif_age, title ="dif_age")
ks.test(datos_combinados$dif_age, "pnorm", mean(datos_combinados$dif_age, na.rm = TRUE), sd(datos_combinados$dif_age, na.rm = TRUE))
```
dif_age does not follow a normal distribution. 


```{r}
ggqqplot(datos_combinados$dif_rank, title = "dif_rank")
ks.test(datos_combinados$dif_rank, "pnorm", mean(datos_combinados$dif_rank, na.rm = TRUE), sd(datos_combinados$dif_rank, na.rm = TRUE))
```
dif_rank does not follow a normal distribution.

```{r}
ggqqplot(datos_combinados$winner_age, title = "winner_age")
ks.test(datos_combinados$winner_age, "pnorm", mean(datos_combinados$winner_age, na.rm = TRUE), sd(datos_combinados$winner_age, na.rm = TRUE))
```
winner_age does not follow a normal distribution.


```{r}
ggqqplot(datos_combinados$loser_age, title = "loser_age")
ks.test(datos_combinados$loser_age, "pnorm", mean(datos_combinados$loser_age, na.rm = TRUE), sd(datos_combinados$loser_age, na.rm = TRUE))
```
loser_age does not follow a normal distribution

```{r}
ggqqplot(datos_combinados$winner_rank, title = "winner_rank")
ks.test(datos_combinados$winner_rank, "pnorm", mean(datos_combinados$winner_rank, na.rm = TRUE), sd(datos_combinados$winner_rank, na.rm = TRUE))
```
winner_rank does not follow a normal distribution

```{r}
ggqqplot(datos_combinados$loser_rank, title = "loser_rank")
ks.test(datos_combinados$loser_rank, "pnorm", mean(datos_combinados$loser_rank, na.rm = TRUE), sd(datos_combinados$loser_rank, na.rm = TRUE))
```

loser_rank does not follow a normal distribution
```{r}
ggqqplot(datos_combinados$best_of, title = "best_of")
ks.test(datos_combinados$best_of, "pnorm", mean(datos_combinados$best_of, na.rm = TRUE), sd(datos_combinados$best_of, na.rm = TRUE))
```
best_of does not follow a normal distribution


```{r}
ggqqplot(datos_combinados$games, title = "games")

ks.test(datos_combinados$games, "pnorm", mean(datos_combinados$games, na.rm = TRUE), sd(datos_combinados$games, na.rm = TRUE))

```

games does not follow a normal distribution.

## 5. Missing data


```{r}
missing_per_column <- colSums(is.na(datos_combinados))

missing_percentage <- missing_per_column * 100 / nrow(datos_combinados) 
missing_percentage


missing_data <- data.frame(
  variable = names(datos_combinados),
  missing_count = colSums(is.na(datos_combinados)),
  total = nrow(datos_combinados)
)

missing_data$percentage <- missing_percentage

Missingdata <- ggplot(missing_data, aes(x = reorder(variable, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "#73edff") +
  labs(title = "Missing data in each variable", 
       x = "Variable", 
       y = "Missing data (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90))

Missingdata

ggsave(filename = "Missingdata.tif", plot = Missingdata, device = "tiff",  dpi = 300)

```


