---
title: "New database"
author: "Ana Abad"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(dplyr)
library(compareGroups)
```


```{r}
load("ATP_final.RData")
data <- dplyr::select(datos_combinados, -LL, -PR)
```


# 1. New database

Each column represents a player, which can be either the loser or the winner of the match. 
That means that variables related to the match or differences between winners/losers are duplicated.


```{r}


players_data_W <- data %>%
  mutate(player_role = "winner") %>%
  rename(
    ID = winner_id, 
    entry = winner_entry,
    hand = winner_hand,
    ioc = winner_ioc,
    age = winner_age,
    rank = winner_rank,
    LL = LLW,
    PR = PRW) %>%
  dplyr::select(tourney_id, tourney_name, surface, tourney_level, year, ID, entry, hand, ioc, age, rank, 
         dif_age, dif_rank, round, LL, PR, player_role)


players_dataL <- data %>%
      mutate(player_role = "loser") %>%
      rename(
        ID = loser_id,
        entry = loser_entry,
        hand = loser_hand,
        ioc = loser_ioc,
        age = loser_age,
        rank = loser_rank,
        LL = LLL,
        PR = PRL
      ) %>%
      dplyr::select(tourney_id, tourney_name, surface, tourney_level, year, ID, entry, hand, ioc, age, rank, 
             dif_age, dif_rank, round, LL, PR, player_role)

players_data <- rbind(players_dataL, players_data_W)


```

# 2. Summary 

```{r}
summary(players_data)
```

# 3. Victories and defeats


## Number of tournaments included in the dataset 
```{r}
tour <- length(unique(players_data$tourney_id))

tour
```

## PR in tournaments

```{r}
players_data %>%
  group_by(tourney_id) %>%
  filter(PR == 1) %>%
  ungroup()
```
## LL matches won and lost 

```{r}

# Filter the data for losing players who were lucky losers
LL_losers <- players_data %>%
  filter(player_role == "loser" & LL == 1) %>%
  # Group by tournament ID and count occurrences
  group_by(tourney_id, tourney_name) %>%
  summarise(matches_lost = n()) %>%
  ungroup()



LL_winners <- players_data %>%
  filter(player_role == "winner" & LL == 1) %>%
  group_by(tourney_id, tourney_name) %>%
  summarise(matches_won = n()) %>%
  ungroup()

played_ll <- players_data %>%
  filter(LL == 1) %>%
  group_by(tourney_id, tourney_name) %>%
  summarise(matches_played = n()) %>%
  ungroup()


LL <- merge(LL_losers, played_ll)
LL$defeats <- LL$matches_lost/LL$matches_played
LL

LLW <- merge(LL_winners, played_ll)
LLW$victories <- LLW$matches_won/LLW$matches_played
LLW


new_ll <- merge(LL, LLW)


```



## PR matches won and lost 

```{r}


PR_losers <- players_data %>%
  filter(player_role == "loser" & PR == 1) %>%
  group_by(tourney_id, tourney_name) %>%
  summarise(matches_lost = n()) %>%
  ungroup()



PR_winners <- players_data %>%
  filter(player_role == "winner" & PR == 1) %>%
  group_by(tourney_id, tourney_name) %>%
  summarise(matches_won = n()) %>%
  ungroup()

played_PR <- players_data %>%
  filter(PR == 1) %>%
  group_by(tourney_id, tourney_name) %>%
  summarise(matches_played = n()) %>%
  ungroup()


PR <- merge(PR_losers, played_PR)
PR$defeats <- PR$matches_lost/PR$matches_played
PR

PRW <- merge(PR_winners, played_PR)
PRW$victories <- PRW$matches_won/PRW$matches_played
PRW

new_PR <- merge(PR, PRW)




```

## Mean victories

```{r}
mean(LLW$victories)
mean(PRW$victories)


## LL victories
binom.test(sum(new_ll$matches_won), sum(new_ll$matches_played))
## LL defeats
binom.test(sum(new_ll$matches_lost), sum(new_ll$matches_played))


mean(LL$defeats)
mean(PR$defeats)



## PR victories
binom.test(sum(new_PR$matches_won), sum(new_PR$matches_played))
## PR defeats
binom.test(sum(new_PR$matches_lost), sum(new_PR$matches_played))




```



# n LL in 1973

```{r}
unique(players_data %>%
         filter(LL == 1 & tourney_id == "1973-540"))
```




```{r}

compareGroups(player_role ~ LL + PR, data = players_data)
```

## Final winners

```{r}

winners <- players_data_W %>%
  filter(round == "F")

winners$entry <- as.character(winners$entry)
sum(winners$entry == "PR", na.rm = TRUE)
sum(winners$entry == "LL", na.rm = TRUE)

```





